---
editor_options: 
  chunk_output_type: console
---

# Build reads data to networks pipeline

```{r}
# load libraries
library(readr)
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(pondr)

# for plotting
library(ggplot2)
library(patchwork)
```

Read file names in data folder.

```{r}
# get data folders
folders = list.dirs("data", recursive = F)

# list csv files for each system, ie, each folder
files = lapply(folders, function(fl) {
  list.files(fl, pattern = "CSV", full.names = TRUE)  
})

# get node data by reading excel files from each folder
# the read excel function is applied to a flexibly constructed
# file path
fish_data = Map(folders, seq(length(folders)),
                f = function(fl, number) {
  read_excel(
    path = sprintf("%s/Data System %i.xlsx", fl, number)
  )
})
```

```{r}
# read data and rename columns
data = Map(
  files, seq(length(files)),
  f = function(fl, num) {
    lapply(fl, read_delim, delim = ";") |>
      bind_rows() |> 
      mutate(
        system = num
      )
  })

data = lapply(data, function(df) {
  rename(
    df,
    date = Date, time = Time, antenna = `Unit number`,
    id = `Transponder code`
  )
})

# clean the data using a pondr function
data = Map(data, fish_data, f = function(reads, nodes) {
  pondr::pr_clean_data(reads, block_ref_df = nodes)
})

# add system id
data = Map(
  data, seq(length(data)),
  f = function(df, num) {
    df$system = num
    df
  }
)
```

It is very important to clean the data before making networks, as the network construction fails when it finds nodes (fish) in the edge list, that are not in the node list. This happens when there are errors in tag reading, introducing a new tag code into the edgelist that is not among the tags in the node data.

```{r}
# bind columns and split by date
data_nt = bind_rows(data) %>%
  mutate(
    date = date(time)
  ) %>%
  nest_by(
    date, system
  )

# add colum for id data
data_nt = mutate(
  data_nt,
  node_data = list(
    fish_data[[system]]
  )
)

# get an example daily network, time bin is 5 seconds by default
networks = Map(
  data_nt$data, data_nt$node_data,
  f = function(reads, nodes) {
    pr_make_network(
      block_df = reads,
      block_ref_df = nodes,
      time_bin = "5 minutes"
    )
  }
)

# save networks
saveRDS(
  networks,
  file = "data/data_networks.Rds"
)
```

```{r}
# make a custom layout that is fixed between plots
# this must be based on an initial network - choose one
data_nt$networks = networks

layout_1 = igraph::layout.kamada.kawai(data_nt$networks[[1]])
layout_2 = igraph::layout.kamada.kawai(data_nt$networks[[2]])

data_nt = 
  mutate(
    data_nt,
    layout = lapply(system, function(x) {
      if(x == 1) layout_1 else layout_2
    })
  )

# make plots and wrap
plots_nt = Map(
    data_nt$networks, data_nt$layout, data_nt$date,
    f = function(n, l, d) {
      pondr::pr_plot_network(
        network = n,
        colour_by = length_mm,
        wt_lim = 150,
        layout = l
      ) +
        scico::scale_fill_scico(
          palette = "hawaii"
        )+
        labs(
          title = d,
          fill = "Length (mm)"
        )+
        theme(
          legend.position = "top",
          legend.key.height = unit(2, "mm")
        )
    }
  )

# add to df
data_nt$plots = plots_nt
```

```{r}
# save as Rds object
saveRDS(data_nt, file = "data/data_networks.Rds")
```

### Plot by system and date

```{r}
data_plots = nest(
  data_nt,
  -system
)
```

```{r}
plot_per_system = Map(
  data_plots$system, data_plots$data,
  f = function(d, ndf) {
    
    # make plots for each day for one system at a time
    plots = Map(
      ndf$networks, ndf$layout, ndf$date,
      f = function(n, l, dt) {
        pondr::pr_plot_network(
          network = n,
          colour_by = length_mm,
          layout = l
        ) +
          ggtitle(
            label = dt
          )
      }
    )
    
    # combine plots for this system
    plot_combined = patchwork::wrap_plots(
      plots, ncol = 5,
      guides = "collect"
    ) &
      theme(
        legend.position = "top",
        legend.key.height = unit(2, "mm")
      )
    
    plot_combined
    
  }  
)

# save combined plots
Map(
  plot_per_system, seq(length(plot_per_system)),
  f = function(gg, n) {
    # save figure
    ggsave(
      gg,
      filename = sprintf("figures/fig_daily_sn_system_%i.png", n),
      height = 300, width = 410, units = "mm"
    )
  }
)
```


```{r}
# wrap all plots
figure = wrap_plots(
  nt_plots,
  guides = "collect"
) &
  theme(
    legend.position = "top"
  )

# save figure
ggsave(
  filename = "figures/fig_show_sn.png"
)
```
